---
title: "Homework 1"
author: "Varun Saxena"
format:
  pdf:
    output-file: "saxena-v-hwk1-3"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)

library(knitr)


#install.packages("knitr")
```



```{r}
#| include: false
#| eval: true
 
load("Hwk1_workspace.Rdata")
```


\newpage
# Enrollment Data

\vspace{.2in}
\noindent 1. How many observations exist in your current dataset?<br>

19126783

\newpage
\noindent 2. How many different *plan_types* exist in the data? <br>

There are `r nrow(plan_count_table)` unique plan types that exist in the data.


\newpage
\noindent 3. Provide a table of the count of plans under each plan type in each year.<br>

```{r} 
#| echo: false
#| label: tbl-plans
#| tbl-cap: "Plan types by year"

#options(knitr.kable.NA = 0)
#knitr::kable(plan.type.year1, 
             #col.names=c("Plan Type","2010","2011",
                         #"2012","2013","2014","2015"),
             #format.args=list(big.mark=","), booktabs = TRUE) %>%
             #kable_styling(latex_options=c("scale_down"))
kable(plan_count_table)
```




\newpage
\noindent 4. Remove all special needs plans (SNP), employer group plans (eghp), and all "800-series" plans. Provide an updated table after making these exclusions.


```{r}
#| echo: false
#| label: tbl-plans2
#| tbl-cap: Revised plan types by year

#options(knitr.kable.NA = 0)
#knitr::kable(plan.type.year2, 
             #col.names=c("Plan Type","2010","2011",
                         #"2012","2013","2014","2015"),
             #format.args=list(big.mark=","), booktabs = TRUE) %>%
             #kable_styling(latex_options=c("scale_down"))
kable(plan_count_table_after_exclusions)

``` 




\newpage
\noindent 5. Merge the the contract service area data to the enrollment data and restrict the data only to contracts that are approved in their respective counties. Limit your dataset only to plans with non-missing enrollment data. Provide a graph showing the average number of Medicare Advantage enrollees per county from 2008 to 2015.<br>

I used the inner join function to join the relevant datasets. Then, I used a filter to exclude entries with missing data. The resulting plot contains the graph of Medicare Advantage enrollees.

```{r}
#| echo: false
#| label: fig-enroll
#| fig-cap: "Average Enrollment"

fig.avg.enrollment
```


\newpage
# Premium Data

\noindent 6. Merge the plan characteristics data to the dataset you created in Step 5 above. Provide a graph showing the average premium over time. Don't forget about formatting!<br>

To do this, I merged in the market penetration and state-level data into the existing dataset created in the previous step. From here, I created a graph that demonstrated the premium price over time.


```{r}
#| echo: false
#| label: fig-premium
#| fig-cap: "Average Premiums"

fig.avg.premium
```



\newpage
\noindent 7. Provide a graph showing the percentage of $0 premium plans over time. Also...remember to format things.

Here is the graph of the percentage of $0 premum plans over this time period:


```{r}
#| echo: false
#| label: fig-zero
#| fig-cap: "Share of 0 premium plans"

fig.percent.zero
```

\newpage
\noindent 8. Why did we drop the "800-series" plans?<br>

800-series plans are only available to a select group of individuals. Since most people cannot opt into these types of plans, they should be excluded from the dataset.

\newpage
\noindent 9. Why do so many plans charge a $0 premium? What does that really mean to a beneficiary?<br>

There is a standard premium already paid for Medicare Part B, so this $0 premium is just the excess over that standard premium. Additionally, tax dollars subsidize this program, so that helps to allow these $0 premiums to exist.

\newpage
\noindent 10. Briefly describe your experience working with these data (just a few sentences). Tell me one thing you learned and one thing that really aggravated you.<br>

My experience working with this data involved much trial and error. With little R experience beforehand, I had to rely on my peers, online resources like git, stack overflow, and others. One thing I learned was how to review R documentation to get to the desired outcome. One thing I struggled with was using join functions and adapting my code to Quarto. But now that I have learned this, it should be smoother going forward. Additionally, my graphs did not come out correctly in Quarto, so I would like to remedy that as well.